Emulsions Stabilized by Magnetic Ellipsoidal Particles: A Lattice Boltzmann Study

### 2025-04-12 chapters/LBM_technique.tex
Changes 1. Reformatting for intro, lit review and methods complete

We begin with a restatement of Reynolds Transport Theorem (RTT) which relates the change in a parameter in a volume over time to the change in that quantity in a control volume,
defined as

\begin{equation}
    \frac{d}{dt} \int_V f \, dV = \int_V \frac{\partial f}{\partial t} \, dV + \int_{\partial V} f (\vec{v}_{\text{surf}} \cdot \vec{n}) \, dA
\end{equation}

Where $f$ can be a scalar, vector or tensor, $V$ is the volume of the control element, $t$ is time, $\vec{v}_{surf}$ is the velocity vector of the surface of the volume element,
$\vec{n}$ is the surface normal unit vector. From the RTT, the continuity equation can be written to analyze the mass flux of a control volume given by,

@@

\begin{equation}
    \rho \left(\frac{\partial\vec{u}}{\partial t} + (\vec{u}\cdot\nabla)\vec{u} \right) = -\nabla p + \nabla \cdot \boldsymbol{\tau} + \vec{f}
\end{equation}

where $p$ is the pressure, $\boldsymbol{\tau}$ is the viscous stress tensor, and $\vec{f}$ represents external forces such as gravity or magnetic interactions.

For a Newtonian, isotropic fluid, the viscous stress tensor is given by

\begin{equation}
    \boldsymbol{\tau} = p\mathbf{I} + \eta\left( \nabla \vec{u} + (\nabla \vec{u})^T \right)
\end{equation}

Substituting this stress tensor into the momentum equation yields the incompressible Navier-Stokes equation,

@@

Bijels form through spinodal decomposition of two partially miscible fluids, a process that can be effectively modeled using the Cahn-Hilliard equation, which describes the diffusive dynamics 
of a conserved scalar order parameter $\phi$, typically representing the local concentration difference between the two fluid components,

@@

Here, $M$ is the mobility coefficient and $\mu$ is the chemical potential, which is derived from the variation of the free energy functional $F[\phi]$. This formulation allows for the modeling of 
interfacial tension and the spontaneous formation of bicontinuous domains characteristic of spinodal decomposition.
To capture not only the diffusive separation of phases but also the resulting fluid motion, the Cahn-Hilliard equation is coupled with the Navie-Stokes equations, leading to the 
Cahn-Hilliard-Navier-Stokes (CHNS) framework. This coupling is essential in systems like bijels, where advection play a significant role in domain coarsening and morphology arrest. 
The momentum conservation equation in this framework is written as:

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale = 1]{figures/methods/d3q19_lattice.jpg}
    \caption{D3Q19 lattice demonstrating the rest, nearest and next nearest direction that correspond to the 19 
    directions $(i)$ of the lattice with lattice velocity $c_{i}$. \cite{schmieschek_lb3d_2017} Reproduced from 
    Schmieschek et al. Computer Physics Communications 2017, 217, 149--161, under the Creative Commons license.}
    \label{fig:d3q19_lattice}
\end{figure}

@@

The momentum flux tensor can be further expressed as a function of macroscopic variables,

@@

\subsection{ellipsoidal_particles}
\label{section:ellipsoidal_particles}

>>>>

We begin with a restatement of Reynolds Transport Theorem (RTT) which relates the accumulation of a parameter within a control volume to the difference between the values
of the parameter entering and leaving the volume and the generation and consumption within the volume. The derivation from page 135 to 147 in reference \cite{batchelor_introduction_2000}
is followed. For a non moving and non-deforming volume, the transport equation is given by,

\begin{equation}
    \frac{d}{dt} \int \theta \rho \, dV = \int Q \, dV - \int \theta \rho (\vec{v}_{\text{surf}} \cdot \vec{n}) \, dA
\end{equation}

Where $\theta$ is an intensive quantity of the system such as mass, momentum or energy, $\rho$ is the density of the fluid in the control volume, $Q$ is an effective density
of a source, $V$ is the volume of the control element, $t$ is time, $\vec{v}_{surf}$ is the velocity vector 
of the surface of the volume element and $\vec{n}$ is the surface normal unit vector. \cite{batchelor_introduction_2000} Using the divergence theorem, the expression can be rewritten as

\begin{equation}
    \int \frac{\partial \theta \rho}{\partial t} \, dV = \int Q \, dV - \int \nabla \cdot (\theta \rho \vec{v}_{\text{surf}}) \, dV
\end{equation}

If $\theta = 1$ and $Q = 0$ the continuity equation for fluid density can be written to analyze the mass flux of a control volume given by,

@@

\begin{equation}
    \rho \left(\frac{\partial u_i}{\partial t} + (u_i\cdot\frac{\partial u_i}{x_i}) \right) = \frac{\partial \sigma_{ij}}{\partial x_j} + f_i
\end{equation}

where $f_i$ is the force exerted on the fluid from external forces such as gravity or magnetic interactions and $\sigma_{ij}$ is the stress tensor. For a Newtonian, isotropic fluid, 
the viscous stress tensor is given by

\begin{equation}
    \sigma_{ij} = -p\delta_{ij} + \eta((\frac{\partial u_i}{\partial x_j} + \frac{\partial u_j}{\partial x_i}) - \frac{2}{3}(\frac{\partial u_j}{\partial x_i})^2\delta_{ij})
\end{equation}

Substituting this stress tensor into the momentum equation and using the relation, $\nabla\cdot \vec{u} = 0$ from the incompressibility constraint, the incompressible 
Navier-Stokes equation can be derived which in vector notation is given by,

@@

Bijels form through spinodal decomposition of two partially miscible fluids, a process that can be effectively modeled using the Cahn-Hilliard equation. \cite{cahn_spinodal_1961} While the
original equation is written for a concentration of species $c$. For a binary mixture, it can be recast in terms of a conserved scalar order parameter $\phi$, typically representing the 
local concentration difference between the two fluid components,

@@

Here, $M$ is the mobility coefficient and $\mu$ is the chemical potential, which is derived from the variation of the free energy functional $F[\phi]$ which is the sum of a bulk free energy
$f(\phi)$ and an interfacial contribution. This formulation allows for the modeling of 
interfacial tension and the spontaneous formation of bicontinuous domains characteristic of spinodal decomposition.
To capture not only the diffusive separation of phases but also the resulting fluid motion, the Cahn-Hilliard equation is coupled with the Navier-Stokes equations, leading to the 
Cahn-Hilliard-Navier-Stokes (CHNS) framework. This coupling is essential in systems like bijels, where advection play a significant role in domain coarsening and morphology arrest. 
The momentum conservation equation in this framework is given by,

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale = 1]{figures/methods/d3q19_lattice.jpg}
    \caption{D3Q19 lattice demonstrating the rest, nearest and next nearest direction that correspond to the 19 
    directions $(i)$ of the lattice with lattice velocity $c_{i}$. \cite{schmieschek_lb3d_2017} Reproduced from 
    Schmieschek et al. Computer Physics Communications 2017, 217, 149--161, under the CC-BY license.}
    \label{fig:d3q19_lattice}
\end{figure}

@@

The momentum flux tensor can be further expressed as a sum of the equilibrium and non-equilibrium components of the stress tensor, denoted by 0 and 1 given by,

@@

\subsection{Ellipsoidal_particles}
\label{section:ellipsoidal_particles}
